### 2024-12-04 paper/paper.qmd
update the paper format
update the paper format

---
title: "Understanding School Shootings: Analyzing Trends, Demographics, and Geographic Factors Using Bayesian Methods"
subtitle: "A Data-Driven Investigation into the Predictors of School Shootings in the United States (1999–2024)"
author: 
  - Ruizi Liu
thanks: "Code and data are available at: [https://github.com/RIRI0527/school_shooting_analysis.git](https://github.com/RIRI0527/school_shooting_analysis.git)."
date: today
date-format: long
abstract: "School shootings remain a critical issue in the United States, with far-reaching social and psychological consequences. This study investigates the factors influencing the occurrence of school shootings, focusing on temporal trends, school type, demographic composition, and geographic context. Using a Bayesian logistic regression model, this analysis incorporates both observed and simulated data to explore the relationship between these predictors and the likelihood of a shooting. The results reveal significant associations between the predictors and school shootings. Temporal trends indicate an alarming increase in incidents over time, while public schools and urban locales exhibit higher risks compared to private schools and rural areas. Demographic composition also plays a role, with schools characterized by specific racial majorities showing distinct patterns in the likelihood of shootings. This report emphasizes the need for targeted, data-driven interventions that consider institutional and geographic variations in risk. By leveraging probabilistic modeling, this study provides actionable insights for policymakers aiming to mitigate the occurrence of school shootings and improve the safety of educational environments across the United States. "
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#### Installing the packages ####
required_packages <- c("tidyverse", "here", "maps")
for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, dependencies = TRUE)
    library(p, character.only = TRUE)
  }
}
```

```{r}
#| include: false

#### Loading the packages ####
library(tidyverse)
library(here)
library(maps)
library(ggplot2)
library(knitr)
library(brms)
library(bayesplot)
library(knitr)
library(kableExtra)
```

```{r}
#| include: false

#### Loading the dataset ####
data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
shooting_data <- data %>%
  filter(data_source == "raw")
non_shooting_data <-data %>%
  filter(data_source == "synthetic")
```

# Introduction
School shootings are among the most tragic and impactful events in the United States, leaving lasting scars on communities and prompting urgent calls for effective prevention measures. While public discourse often focuses on the immediate aftermath of such events, it is equally important to examine the underlying factors contributing to their occurrence. This report aims to analyze the temporal, institutional, demographic, and geographic predictors of school shootings, drawing insights from historical data to inform future policies.

This analysis utilizes a Bayesian logistic regression model to explore whether factors such as the year, school type, demographic composition, and urbanicity influence the likelihood of school shootings. Bayesian methods are particularly suitable for this investigation, as they allow for the incorporation of prior knowledge while providing probabilistic interpretations of the results. By using both observed and simulated data, the model offers a comprehensive perspective on the risk factors associated with these incidents.

The findings presented in this report highlight critical trends and disparities that warrant further attention. For example, the temporal trend of increasing school shootings underscores the urgency of targeted interventions. Similarly, the variations in risk between public and private schools, urban and rural areas, and schools with differing demographic compositions suggest that a one-size-fits-all approach may not be sufficient. By identifying and analyzing these patterns, this study seeks to contribute to the broader conversation on preventing school shootings in the United States.

# Data {#sec-data}

## Overview

The dataset originates from the Washington Post [@washingtonpost_school_shootings] and spans 1999 to 2024, documenting school shootings in the United States. The raw dataset includes 416 observations and 50 variables, detailing information such as the time, location, racial demographics of schools, and shooter characteristics. This study extends the dataset by simulating additional observations to balance the binary outcome variable (shooting occurrence), resulting in a dataset with 830 observations and 31 variables. The dataset provides a robust basis for examining the predictors of school shootings, including temporal trends, demographic composition, and urbanicity.

## Measurement

The dataset is a product of rigorous investigation by the Washington Post, combining diverse and credible sources to create a comprehensive record of school shootings from 1999 to the present [@washingtonpost_school_shootings]. Real-world incidents—where a gun was fired and at least one person was injured or killed on school property—were systematically transformed into structured data entries through a transparent and detailed process.

### Data Sources and Process

The measurement process involved triangulating information from:
- **News Reports**: Verified through multiple outlets for accuracy.
- **Law Enforcement Records**: Provided timelines, locations, and key details.
- **School Data**: Publicly available records enriched the dataset with demographic and institutional contexts.

Each incident was identified, verified, and contextualized to ensure accuracy. Variables such as school type, urban locale description, and racial composition were drawn from NCES data and school enrollment records, with additional details cross-referenced for completeness. Where gaps existed, synthetic data were carefully generated to maintain analytical consistency.

### Strengths of the Dataset

The dataset’s strengths lie in its:
- **Comprehensiveness**: Captures a wide range of school shootings with detailed contexts.
- **Transparency**: Relies on well-documented and publicly accessible sources.
- **Relevance**: Includes key variables such as urbanicity and demographics critical to understanding school shootings.

These qualities ensure that the dataset provides a reliable foundation for examining the complex factors contributing to school shootings.

## Data Cleaning

We used the`R`programming language [@citeR], the `here` package [],
the `dplyr` package [@R_dplyr], the `tidyr` package [@R_tidyr], the `maps` package [@R_maps], the `ggplot2` package [@R_ggplot2], and the`scale` package [@R_scales] to clean the data. We used the `tidyverse` package [@tidyverse], the `maps` package [@maps], the `ggplot2` package [@R_ggplot2], the `knitr` package [@knitr], the `kableExtra` package [@kableExtra] to plot the dataset and the visualization.

In doing so, R code was adapted from [@tellingstories].

First we did some general cleaning of this dataset, including replacing NAs in the data with means and changing the form of dates in the data.In addition, because each school had different enrolments, we calculated the percentage of different ethnicities in each school and selected the highest percentage of ethnicities in each school.

In addition, because the original data only recorded the shootings that occurred and not the ones that did not, but we wanted to simulate the logistic regression model, we generated synthetic data based on the cleaned dataset to reflect the data that did not occur. The main components that were simulated were latitude, longitude, school ethnicity, lunch, and staffing, and to ensure that the data were reasonable, the synthetic data were generated within the allowable range of variability (a reasonable float between the maximum and minimum of the variable), and to ensure that the model was implementable, the synthetic data were generated within the allowable range of variation.Synthetic data and raw data both have 415 observations.at the same time, we made a distinction between synthetic data and raw data inside the dataset by labelling their data source in data_source and creating a new binary variable to indicate whether a shooting occurred or not.

Finally, we cleaned the dataset of some unnecessary variables, such as the name and id of the school, the specific city neighbourhoods where the shootings took place (we focused our analysis on those states specifically), the shooter's access to the gun, and so on.

In the end, we cleaned up the data to include 830 observations and 31 variables, the dataset can be find in @tbl-data-overview, only specific columns shows in the table.

```{r}
#| label: tlb-data-overview
#| tib-cap: The specific columns in the dataset
#| echo: false
#| warning: false

# Subset the dataset for demonstration (first 10 rows and selected columns)
summary_data <- data[1:10, c("year", "state", "school_type", "top_1_races", "ulocale_desc", "data_source", "school_shooting")]

# Create the kable table
kable(summary_data, 
      col.names = c("Year", "State", "School Type", "Top 1 Race", "Ulocale", "Data Source", "School Shooting"),
      caption = "Summary of School Shooting Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                position = "left")
```

## Data visualization

```{r}
#| label: fig-school-type
#| fig-cap: Distribution of School Types
#| echo: false
#| warning: false
# Count occurrences of each category in 'school_type'
school_type_counts <- as.data.frame(table(shooting_data$school_type))

# Rename the columns for clarity
colnames(school_type_counts) <- c("school_type", "Count")

# Create a bar plot
ggplot(school_type_counts, aes(x = reorder(school_type, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(
       x = "School Type",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The binary outcome variable indicates whether a school shooting occurred (1) or not (0). As shown in @fig-school-type, public schools dominate the dataset, reflecting their prominence in the U.S. educational landscape. Private schools are significantly underrepresented, necessitating careful modeling to account for this imbalance.

```{r}
#| label: fig-ulocale
#| fig-cap: Distribution of Urban Locale Descriptions
#| echo: false
#| warning: false

filtered_data <- data %>%
  filter(data_source == "raw" & !is.na(ulocale_desc))

ggplot(filtered_data, aes(x = ulocale_desc, fill = ulocale_desc)) +
  geom_bar() +
  geom_text(stat = "count", 
            aes(label = paste0(round(..count../sum(..count..) * 100, 1), "%")), 
            vjust = -0.2, color = "black", size = 3.5) +
  scale_fill_viridis_d(option = "D") + # Use a visually appealing viridis palette
  labs(
    x = "Urban Locale Description",
    y = "Count"
  ) +
  theme_minimal(base_size = 12) + # Set base font size for overall consistency
  theme(
    legend.position = "none", # Remove legend
    plot.title = element_text(face = "bold", hjust = 0.5), # Center-align title
    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis labels for readability
  )
```

@fig-ulocale illustrates the distribution of urban locales for shootings, with **City: Large** and **Suburb: Large locales** representing the majority of incidents. Smaller locales, such as **Town: Fringe** and **Rural: Remote**, are underrepresented. This trend reflects the impact of population density and urbanicity on school shooting occurrences.

```{r}
#| label: fig-state-map
#| fig-cap: Shooting Occurrences by State
#| echo: false
#| warning: false

# Group shooting data by state and count occurrences
state_counts <- shooting_data %>%
  group_by(state) %>%
  summarize(occurrences = n())

# Convert state names to lowercase for consistency
state_counts$state <- tolower(state_counts$state)

# Create a mapping of state abbreviations to full names
state_abbreviation_to_full <- c(
  "al" = "alabama", "ak" = "alaska", "az" = "arizona", "ar" = "arkansas", 
  "ca" = "california", "co" = "colorado", "ct" = "connecticut", "de" = "delaware", 
  "fl" = "florida", "ga" = "georgia", "hi" = "hawaii", "id" = "idaho", 
  "il" = "illinois", "in" = "indiana", "ia" = "iowa", "ks" = "kansas", 
  "ky" = "kentucky", "la" = "louisiana", "me" = "maine", "md" = "maryland", 
  "ma" = "massachusetts", "mi" = "michigan", "mn" = "minnesota", "ms" = "mississippi", 
  "mo" = "missouri", "mt" = "montana", "ne" = "nebraska", "nv" = "nevada", 
  "nh" = "new hampshire", "nj" = "new jersey", "nm" = "new mexico", "ny" = "new york", 
  "nc" = "north carolina", "nd" = "north dakota", "oh" = "ohio", "ok" = "oklahoma", 
  "or" = "oregon", "pa" = "pennsylvania", "ri" = "rhode island", "sc" = "south carolina", 
  "sd" = "south dakota", "tn" = "tennessee", "tx" = "texas", "ut" = "utah", 
  "vt" = "vermont", "va" = "virginia", "wa" = "washington", "wv" = "west virginia", 
  "wi" = "wisconsin", "wy" = "wyoming"
)

# Replace state abbreviations with full names
state_counts$state <- state_abbreviation_to_full[state_counts$state]

# Get map data for U.S. states
us_map <- map_data("state")

# Merge state data with map data
us_map_data <- left_join(us_map, state_counts, by = c("region" = "state"))

# Replace NA occurrences with 0 for states with no shootings
us_map_data$occurrences[is.na(us_map_data$occurrences)] <- 0

# Add a data frame of state centers for labeling
state_centers <- us_map %>%
  group_by(region) %>%
  summarize(
    long = mean(range(long)),
    lat = mean(range(lat))
  ) %>%
  mutate(state_abbreviation = toupper(names(state_abbreviation_to_full)[match(region, state_abbreviation_to_full)]))  # Get state abbreviations

# Plot the choropleth map with customized theme
ggplot(us_map_data, aes(x = long, y = lat, group = group, fill = occurrences)) +
  geom_polygon(color = "white") +
  geom_text(data = state_centers, aes(x = long, y = lat, label = state_abbreviation), 
            inherit.aes = FALSE, size = 3, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "gray80") +
  theme_void() +  # Removes all axes, grid lines, and background
  theme(
    legend.position = "right",  # Keeps the legend in its default position
    plot.title = element_text(hjust = 0.5, size = 16)  # Center-align and adjust the title size
  ) +
  labs(
    fill = "Occurrences"
  )
```

@fig-state-map maps the distribution of shootings across states, highlighting regions like California, Texas, and Florida as hotspots due to their large populations and extensive school systems.

```{r}
#| label: fig-year
#| fig-cap: Number of Shootings by Year
#| echo: false
#| warning: false

year_counts <- as.data.frame(table(data$year))

# Rename columns for better understanding
colnames(year_counts) <- c("Year", "Count")

# Plot the data
ggplot(year_counts, aes(x = as.numeric(as.character(Year)), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
       x = "Year",
       y = "Number of Shootings") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@fig-year depicts an alarming increase in school shootings over time. This trend emphasizes the urgency of understanding and mitigating factors contributing to these incidents.

# Model

## Model set-up

To analyze the factors contributing to school shootings, we use a Bayesian logistic regression model. This approach is well-suited to modeling the binary outcome variable (whether a school shooting occurred) while incorporating prior knowledge about the predictors. The model is defined as:

$$
y_i \mid \pi_i \sim \text{Bernoulli}(\pi_i)
$$

$$
\text{logit}(\pi_i) = \beta_0 + \beta_1 \cdot \text{year}_i + \beta_2 \cdot \text{school\_type}_i + \beta_3 \cdot \text{ulocale\_desc}_i + \beta_4 \cdot \text{enrollment}_i + \beta_5 \cdot \text{top\_1\_races}_i
$$

where:

- $y_i$: Binary response variable indicating whether a school shooting occurred (\(y_i = 1\)) or not (\(y_i = 0\)).

- $\pi_i$: Probability of a school shooting for observation \(i\).

- $\beta_k$: Coefficients representing the influence of predictors on the log-odds of a school shooting.

### 3.2 Variables and Priors

#### Predictors
- **Year**: Captures temporal trends in shootings.
- **School Type**: Public vs. private schools.
- **Urban Locale**: Urbanicity levels from NCES data.
- **Enrollment**: Total number of students in the school.
- **Dominant Race**: Primary racial group in the school.

#### Priors
- Intercept $\beta_0\$: $\text{Normal}(0, 5)$, reflecting baseline uncertainty.
- Coefficients $\beta_k$: $\text{Normal}(0, 2.5)$, allowing moderate variability.

These priors prevent overfitting while allowing data-driven insights.

## Results

### Model Estimates

Table @tbl-model-summary presents the Bayesian logistic regression coefficients, while Figure @fig-credible-intervals visualizes 90% credible intervals for the predictors.

```{r tbl-model-summary, echo=FALSE, warning=FALSE, startup_message = FALSE}

#| fig-cap: Bayesian Logistic Regression Coefficients

# config_modelsummary(startup_message = FALSE)
library(modelsummary)
model <- readRDS(file = here::here("models/model_bayes.rds"))

modelsummary(model, 
             output = "kableExtra",
             stars = FALSE,
             statistic = "std.error",
             title = "Bayesian Logistic Regression Coefficients")
```

Key findings include:

\• Temporal Trends: Positive association with increasing years.

\• School Type: Higher risk in public schools.

\• Urban Locale: Elevated risk in urban and suburban schools.

\• Enrollment: Modest increase in risk with larger enrollment.

\•Dominant Race: Schools with predominantly White or Hispanic populations are at higher risk.

### Posterior Predictive Checks

Figure @fig-ppc shows the posterior predictive checks (PPCs), which validate the model's fit by comparing observed and simulated data.

## Parameter Estimates

The parameter estimates from the Bayesian logistic regression model are shown in @fig-credible-intervals.
Positive coefficients indicate an increased likelihood of school shootings, while negative coefficients indicate a decreased likelihood.

The results highlight several significant effects:

-   **Temporal Trends**: The coefficient for `year` was positive ($\beta = 0.065$, 95% CI: \[0.042, 0.088\]), suggesting an increasing trend in school shootings over time.

-   **School Type**: Public schools were associated with higher odds of shootings compared to private schools ($\beta = 0.788$, 95% CI: \[0.069, 1.523\]).

-   **Demographic Composition**: Schools with a majority of Asian ($\beta = -6.007$, 95% CI: \[-13.346, -0.230\]) or Black students ($\beta = -5.194$, 95% CI: \[-12.128, -0.255\]) showed significantly reduced odds of school shootings.

-   **Urbanicity**: Schools in rural and suburban areas had reduced odds of shootings compared to urban schools:
    -   Remote rural areas ($\beta = -3.162$, 95% CI: \[-4.168, -2.229\])
    -   Mid-sized suburban areas ($\beta = -2.528$, 95% CI: \[-3.314, -1.792\])

These findings suggest temporal, institutional, demographic, and geographic factors all contribute to the likelihood of school shootings.

## Diagnostics

Model diagnostics confirmed the robustness of the Bayesian logistic regression.

Key diagnostic plots include:

1.  **Posterior Predictive Check** @fig-ppc: The posterior predictive distribution aligns closely with the observed data, indicating a good model fit.

2.  **Convergence**: @fig-trace-plots displays trace plots for key parameters, showing well-mixed chains that reached stationarity. All parameters had ($\hat{R} < 1.1$), as shown in @fig-rhat.

3.  **Credibility Intervals** @fig-credible-intervals: The $90%$ credibility intervals for each parameter are plotted, highlighting significant predictors (intervals that exclude 0).

These diagnostics provide strong evidence of model convergence and fit, supporting the reliability of the results.

## Implications

The results have several implications for understanding and addressing school shootings:

- The increasing trend over time underscores the urgency of targeted interventions to reduce school shootings.

- The higher risk in public schools suggests that policies and resources should prioritize these institutions.

- Geographic patterns indicate that rural and suburban schools have lower risks compared to urban schools, which may reflect differences in community or school characteristics.

- Demographic findings highlight the need for equitable and inclusive approaches to address structural inequalities potentially influencing school shootings.

Overall, these findings provide a robust framework for designing policies to mitigate the risk of school shootings.

# Discussion

This study examined the temporal, institutional, demographic, and geographic factors influencing the likelihood of school shootings using a Bayesian logistic regression model. The model provided robust insights, with credible intervals indicating the degree of uncertainty around the estimated effects @tbl-modelresults.

## Key Findings

The results revealed several significant predictors of school shootings:

-   **Temporal Trends**: The positive association between `year` and the likelihood of school shootings highlights a concerning trend over time.
    This finding underscores the urgency of implementing effective preventative measures to reverse this trajectory.

-   **Institutional Context**: Public schools showed a significantly higher risk compared to private schools.
    This result suggests the need for targeted interventions in public school settings, potentially focusing on resource allocation, security measures, and community engagement.

-   **Demographic Composition**: The negative coefficients for schools with predominantly Asian or Black populations indicate a reduced likelihood of shootings in these contexts.

    These results call for further investigation into how demographic, cultural, or social factors may create protective environments.

-   **Geographical Variation**: Rural and suburban schools were consistently associated with lower odds of school shootings compared to urban schools.
    This pattern emphasizes the importance of location-specific prevention strategies that account for urban challenges, such as population density and resource distribution.

## Strengths and Limitations

### Strengths:

1.  **Bayesian Framework**: The use of Bayesian methods allowed for the incorporation of prior information and the quantification of uncertainty, resulting in more nuanced interpretations of the model coefficients.

2.  **Model Diagnostics**: Comprehensive diagnostic checks ([Figures 1-5](#fig-ppc)) confirmed model convergence and goodness-of-fit, lending credibility to the findings.

### Limitations:

1.  **Generalizability**: While the model captures key predictors, external factors such as policy changes or economic conditions were not included. These factors may mediate or moderate the observed relationships.

2.  **Data Availability**: The quality and completeness of the data used to fit the model may have introduced bias or omitted variable effects. Future studies should aim to include additional covariates to improve explanatory power.

## Implications and Future Directions

The findings have several implications for policy and practice:

- **Policy Interventions**: Addressing the rising trend in school shootings will require concerted efforts at the institutional level, including targeted investments in public schools.

- **Equity and Inclusion**: The protective effects associated with certain demographic groups suggest that inclusive, community-driven strategies may play a role in prevention.

- **Location-Specific Strategies**: Policymakers should tailor interventions to the unique needs of urban, suburban, and rural schools, recognizing the contextual factors that influence risk.

Future research should explore additional variables, such as socioeconomic factors, legislative impacts, and mental health resources, to provide a more comprehensive understanding of the dynamics driving school shootings.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of the Bayesian Logistic Regression Model"

# Load the model
model <- readRDS(file = here::here("models/model_bayes.rds"))

# Extract coefficients and credible intervals
summary_model <- summary(model)
coefficients <- as.data.frame(summary_model$fixed)
coefficients <- coefficients[, c("Estimate", "l-95% CI", "u-95% CI")]

# Replace NA values with "N/A"
coefficients[is.na(coefficients)] <- "N/A"

# Format row names
rownames(coefficients) <- gsub("b_", "", rownames(coefficients))

# Create a clean table
knitr::kable(
  coefficients,
  format = "html", # Change to "latex" for PDF output or "markdown" for plain Markdown
  caption = "Summary of the Bayesian Logistic Regression Model",
  col.names = c("Estimate", "Lower CI", "Upper CI"),
  digits = 3
)
```

\newpage

# Appendix

## A Additional Data Details

This analysis relied on data from the Washington Post dataset, which provides detailed records of school shootings in the United States from 1999 to 2024. To enhance the robustness of the model, synthetic data were generated for schools where shootings did not occur, following the methodology described in [Section 2.3](#data-cleaning). The inclusion of synthetic data allowed for a balanced dataset of 830 observations and 31 variables.

The data cleaning process included:
- Replacing missing values with means.
- Converting date formats for consistency.
- Calculating demographic percentages for each school and retaining only the most prominent demographic group per school.
- Excluding variables that were not relevant to the analysis, such as school names and shooter-specific details.

@tbl-modelresults provides a summary of the cleaned dataset and key variables.

## B Model Details

### B.1 Posterior Predictive Check

The posterior predictive check was performed to evaluate how well the model captures the observed data distribution. @fig-ppc demonstrates that the model predictions align closely with the observed data, indicating a good model fit.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-ppc
#| fig-cap: "Posterior predictive check"

pp_check(model, type = "dens_overlay") +
  theme_classic() +
  theme(legend.position = "bottom")
```

### B.2 Diagnostics

To ensure the reliability of the Bayesian logistic regression model, the following diagnostics were conducted:
1. **Trace Plots:** @fig-trace-plots shows well-mixed chains for all parameters, confirming convergence.
2. **Rhat Plot:** @fig-rhat illustrates that all parameters have $\hat{R} \leq 1.1$, indicating no convergence issues.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-trace-plots
#| fig-cap: "Trace plots for selected parameters"

mcmc_trace(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                     "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  theme_classic() +
  ggtitle("Trace Plots")
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat diagnostic plot for model convergence"

# Extract Rhat values using built-in brms diagnostics
rhat_vals <- brms::rhat(model)

# Visualize Rhat values
bayesplot::mcmc_rhat(rhat_vals) +
  theme_classic() +
  ggtitle("Rhat Diagnostic Plot")
```
### B.3 Comparison of Priors and Posteriors

@fig-posterior-vs-prior compares the posterior distributions with the priors, highlighting how the data influenced the parameter estimates.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-posterior-vs-prior
#| fig-cap: "Comparison of posterior and prior distributions"

mcmc_areas(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                     "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  ggtitle("Posterior Distributions") +
  theme_classic()
```

## C Survey and Sampling Methodology

### C.1 Population, Frame, and Sample
- **Population:** All K-12 schools in the United States.
- **Frame:** Schools from the Washington Post dataset, combined with synthetic data for schools without shootings.
- **Sample:** Stratified by school type, urban locale description, and dominant demographic group.

### C.2 Sampling and Recruitment
- Synthetic data were generated within observed variable ranges to represent schools where shootings did not occur.
- Observed data were supplemented by demographic, geographic, and enrollment metrics.

### C.3 Questionnaire and Validation
- **Survey Focus:** School security measures, demographic composition, geographic details, and historical incidents.
- **Validation:** Cross-referenced data with administrative records to ensure accuracy.

## D Simulation Details

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-occurrence-map
#| fig-cap: School Shooting Occurrence

# Load the base map of the USA
usa_map <- map_data("state")

# Plot the map with shooting locations
ggplot() +
  geom_polygon(data = usa_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = NA) +
  geom_point(data = shooting_data, aes(x = long, y = lat),
             color = "red", alpha = 0.6, size = 2) +
  labs(
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

```
To simulate the occurrence of school shootings, synthetic data were generated with random variability:
- Geographic coordinates were adjusted within observed ranges.
- Demographics and enrollment were assigned proportionally based on observed distributions.

@fig-occurrence-map visualizes the geographic distribution of shooting occurrences, highlighting both observed and synthetic data points.
## E Implications of Results

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-credible-intervals
#| fig-cap: "90% credibility intervals for selected predictors"

mcmc_intervals(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                         "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  theme_classic() +
  ggtitle("90% Credibility Intervals")

```

The analysis revealed the following key insights:
1. **Temporal Trends:** The increasing trend in shootings over time (@fig-trend) underscores the need for immediate interventions.

2. **Institutional and Demographic Context:** Public schools and urban locales were associated with higher risks, as shown in @fig-credible-intervals.
These findings provide actionable insights for policymakers aiming to mitigate risks associated with school shootings.

\newpage

# References